---
title: "Filmexplore"
author: "mike"
date: "26 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




catorgoical vars for total and weekend gross
```{r}

fvars<-Film1[,c(2,10:12)]
summary(fvars)
studiotab=table(Film$Studio)

genretab=table(Film$Genre)
ratingtab=table(Film$Rating)
seq.oritable=table(Film$Seq.Ori)

barplot(studiotab)
barplot(genretab)
barplot(ratingtab)
barplot(seq.oritable)

mosaicplot(xtabs(~Film$Genre+Film$Rating))
Film1$Genre[which(Film1$Rating=="G")]

boxplot(Film1$Awd~Film1$Season)

boxplot(Film1$Nom~Film1$Season)
```


NAs


```{r}
#145,845
#189,398,758,885
load("FilmData.Rda")

Film$Meta[c(145,189,398,758,845,885)]=NA

save(Film,file="Film.Rda")
library(Hmisc)


metasubs=aregImpute(as.formula(paste('~',paste(names(Film)[-1],collapse = "+"))),data=Film)
subbox<-metasubs$imputed$Meta
save(subbox,file="subbox.Rda")
load("Filmomit.Rda")

plot(Film.omit$Imdb,Film.omit$Meta)
matpoints(Film$Imdb[c(145,189,398,758,845,885)],metasubs$imputed$Meta)

#building 5 data frames
Film1<-Film
Film2<-Film
Film3<-Film
Film4<-Film
Film5<-Film

Film1$Meta[c(145,189,398,758,845,885)]<-metasubs$imputed$Meta[,1]
Film2$Meta[c(145,189,398,758,845,885)]<-metasubs$imputed$Meta[,2]
Film3$Meta[c(145,189,398,758,845,885)]<-metasubs$imputed$Meta[,3]
Film4$Meta[c(145,189,398,758,845,885)]<-metasubs$imputed$Meta[,4]
Film5$Meta[c(145,189,398,758,845,885)]<-metasubs$imputed$Meta[,5]

save(Film1,file="Film1.Rda")
save(Film2,file="Film2.Rda")
save(Film3,file="Film3.Rda")
save(Film4,file="Film4.Rda")
save(Film5,file="Film5.Rda")


load("Film1.Rda")

library(ggplot2)
gplot1=ggplot(Film1)
gplot2=ggplot(Film2)

#meta imdb corr
gplot1+geom_point(aes(x=Imdb,y=Meta,color=log(Imdbvotes)))+geom_smooth(aes(x=Imdb,y=Meta),method="loess",se=FALSE,col="red")


#theatres corr
gplot1+geom_point(aes(x=NumOpenTheatres,y=NumTotalTheatres,color=log(Nom)))
gplot1+geom_point(aes(x=jitter(NumOpenTheatres),y=jitter(NumTotalTheatres),color=Season))
```

ggplot

```{r}

library(ggplot2)
gplot1=ggplot(Film1)
gplot2=ggplot(Film1)
```

```{r}
cvars<-Film1[,c(3:6,9,13:17)]
x.cor=round(cor(cvars),2)
#pairs(cvars)

#meta imdb corr
x.cor["Imdb","Meta"]
gplot2+geom_point(aes(x=Imdb,y=Meta,color=log(Budget)))+geom_smooth(aes(x=Imdb,y=Meta),method="loess",se=FALSE,col="red")


#theatres corr
x.cor["NumOpenTheatres","NumTotalTheatres"]
gplot2+geom_point(aes(x=NumOpenTheatres,y=NumTotalTheatres,color=Awd))


#nom awd corr
x.cor["Nom","Awd"]
gplot2+geom_point(aes(x=Nom,y=Awd,color=Meta))+geom_smooth(aes(x=Nom,y=Awd),method="loess",se=FALSE,col="red")

```

transformations
```{r}
library(boot)
Film1$MetaT<-logit(Film1$Meta/100)
Film1$ImdbT<-logit(Film1$Imdb/10)


```


response=opening weekend
formula making
```{r}
#nom,meta
form1=as.formula(paste("log(Opening.weekend)~",paste(names(Film1)[c(2,6,7,8,9,10,11,12,22)],collapse = "+")))

#nom, imdb
form2=as.formula(paste("log(Opening.weekend)~",paste(names(Film1)[c(2,6,7,8,9,10,11,12,23)],collapse = "+")))


```


response=postgross
formula making
```{r}
#nom,meta
form1pg=as.formula(paste("log(postopengross)~",paste(names(Film1)[c(2,4,5,7,8,9,10,11,12,22)],collapse = "+")))

#nom, imdb
form2pg=as.formula(paste("log(postopengross)~",paste(names(Film1)[c(2,4,7,8,9,10,11,12,23)],collapse = "+")))



```

lm weekend gross
```{r}

#aic1=rep(0,2)

lm1=glm(form1,gaussian,Film1)
lm1.s=step(lm1)
summary(lm1.s)

#aic1[1]=AIC(lm1.s)


```

```{r}
lm1=glm(form2,gaussian,Film1)
lm1.s.2=step(lm1)
summary(lm1.s.2)

aic1[2]=AIC(lm1.s.2)
```




```{r}
aic1 #form1/lm1.s is best
plot(lm1.s)
cv.glm(Film1,lm1.s)$delta

```

improve
```{r}
lm1i=glm(formula = log(Opening.weekend) ~ Studio + NumOpenTheatres + 
    Year + Season + Budget + Seq.Ori + MetaT, family = gaussian, 
    data = Film1)

cv.glm(Film1,lm1i)$delta
plot(lm1i)
summary(lm1i)
```



glm weekend gross

```{r}
aic2=rep(0,2)

glm1=glm(form1,family = Gamma(identity),data=Film1)
glm1.s=step(glm1)
summary(glm1.s)

aic2[1]=AIC(glm1.s)

```

```{r}

glm1=glm(form2,family = Gamma(identity),data=Film1)
glm1.s.2=step(glm1)
summary(glm1.s.2)

aic2[2]=AIC(glm1.s.2)

```



```{r}
aic2#form 1, glm1.s is best
plot(glm1.s)
cv.glm(Film1,glm1.s)$delta


```


```{r}
glm1i=glm(formula = log(Opening.weekend) ~ Studio + NumOpenTheatres + 
    Year + Season + Budget + Seq.Ori + MetaT, family = Gamma(identity), 
    data = Film1)

cv.glm(Film1,glm1i)$delta
plot(glm1i)
summary(glm1i)
```

lm post gross

```{r}
which(Film1$postopengross==0)#912
aic3=rep(0,2)

lm2=glm(form1pg,gaussian,Film1,subset=-912)
lm2.s=step(lm2)
summary(lm2.s)
aic3[1]=AIC(lm2.s)

```

```{r}
lm2=glm(form2pg,gaussian,Film1,subset=-912)
lm2.s.2=step(lm2)
summary(lm2.s.2)
#aic3[2]=AIC(lm2.s.2)
```



```{r}
aic3 #form2 or lm2.s.2 is best
plot(lm2.s.2)
cv.glm(Film1[-912,],lm2.s.2)$delta

```

improve
```{r}
lm2i=glm(formula = log(postopengross) ~ Studio + NumTotalTheatres  + Year + Opening.weekend + I(Budget^2) + Season + Budget + Genre + ImdbT, family = gaussian, data = Film1, subset = -912)

plot(lm2i)
summary(lm2i)
```



glm postgross
```{r}
aic4=rep(0,2)

glm2=glm(form1pg,family = Gamma(identity),data=Film1,subset = -912)
glm2.s=step(glm2)
summary(glm2.s)
aic4[1]=AIC(glm2.s)
```

```{r}

glm2=glm(form2pg,family = Gamma(identity),data=Film1,subset = -912)
glm2.s.2=step(glm2)
summary(glm2.s.2)
aic4[2]=AIC(glm2.s.2)
```



```{r}
aic4#form2 is best, glm2.s.2
plot(glm2.s.2)
cv.glm(Film1[-912,],glm2.s.2)$delta
```

 weekend gross per theatre

```{r}

#nom,meta
form1pt=as.formula(paste("log(Opening.weekend)~offset(log(NumOpenTheatres))+",paste(names(Film1)[c(2,7,8,9,10,11,12,22)],collapse = "+")))

#nom, imdb
form2pt=as.formula(paste("log(Opening.weekend)~offset(log(NumOpenTheatres))+",paste(names(Film1)[c(2,7,8,9,10,11,12,23)],collapse = "+")))



```

lm per theatre 

```{r}
aic5=rep(0,2)

lm3=glm(form1pt,gaussian,Film1)
lm3.s=step(lm3)
summary(lm3.s)
aic5[1]=AIC(lm3.s)
```

```{r}
lm3=glm(form2pt,gaussian,Film1)
lm3.s.2=step(lm3)
summary(lm3.s.2)
aic5[2]=AIC(lm3.s.2)
```



```{r}
aic5 #form1 is best lm3.s
plot(lm3.s)
cv.glm(Film1,lm3.s)$delta

```


improve

```{r}
lm3i=glm(formula = log(Opening.weekend) ~ Studio + Season + Budget + Seq.Ori + MetaT + offset(log(NumOpenTheatres)), 
    family = gaussian, data = Film1)
summary(lm3i)
plot(lm3i)
cv.glm(Film1,lm3i)
```



glm

```{r}
aic6=rep(0,2)

glm3=glm(form1pt,Gamma(identity),data=Film1)
glm3.s=step(glm3)
summary(glm3.s)
aic6[1]=AIC(glm3.s)
```

```{r}
glm3=glm(form2pt,Gamma(identity),data=Film1)
glm3.s.2=step(glm3)
summary(glm3.s.2)
aic6[2]=AIC(glm3.s.2)
```



```{r}
aic6 #form1
plot(glm3.s)
cv.glm(Film1,glm3.s)$delta
```


improve

```{r}
glm3i=glm(glm3.s$formula,Gamma(identity),data=Film1,subset=-1002)
summary(glm3i)

glm3i2=glm(log(Opening.weekend) ~ Studio + Season + Budget + Seq.Ori + 
    MetaT + offset(log(NumOpenTheatres)),Gamma(identity),data=Film1,subset=-1002)
summary(glm3i2)
plot(glm3i2)

cv.glm(Film1[-1002,],glm3i2)$delta
```


post gross per theatre

```{r}

#nom,meta
form1pgpt=as.formula(paste("log(postopengross)~offset(log(NumTotalTheatres))+",paste(names(Film1)[c(2,5,7,8,9,10,11,12,22)],collapse = "+")))

#nom, imdb
form2pgpt=as.formula(paste("log(postopengross)~offset(log(NumTotalTheatres))+",paste(names(Film1)[c(2,5,7,8,9,10,11,12,23)],collapse = "+")))


```

lm post gross pt

```{r}
aic7=rep(0,2)
lm4=glm(form1pgpt,gaussian,Film1,subset=-912)
lm4.s=step(lm4)
summary(lm4.s)
aic7[1]=AIC(lm4.s)

```

```{r}
lm4=glm(form2pgpt,gaussian,Film1,subset=-912)
lm4.s.2=step(lm4)
summary(lm4.s.2)
aic7[2]=AIC(lm4.s.2)

```



```{r}
aic7 #form2
plot(lm4.s.2)
cv.glm(Film1[-912,],lm4.s.4)$delta
```

improve

```{r}
lm4i=glm(log(postopengross) ~ Studio + Opening.weekend +owi + Year + Season + Genre + ImdbT + offset(log(NumTotalTheatres)),family=gaussian, Film1,subset=-912)
plot(lm4i)
summary(lm4i)

```

```{r}
ex=which(predict(lm4.s.2)>19.5)
ex=c(1,23 ,24 ,25  ,43  ,44,  45,  46, 127, 148, 149, 150, 151, 211, 277, 279, 280, 339, 401, 402, 404, 405, 461, 463, 502, 525, 526, 589, 590, 591, 651, 652, 717, 760, 783, 784, 787, 847, 913, 915, 916, 917, 979, 980)
Film1$Film[ex]
Film1$Seq.Ori[ex]
Film1$Budget[ex]
hist(Film1$Opening.weekend[ex])
hist(Film1$Opening.weekend[-ex])
Film1$NumTotalTheatres[ex]
Film1$Year[ex]
Film1$Season[ex]
Film1$owi=cut(Film1$Opening.weekend,c(0,75000000,250000000))
```



glm postgross pt
```{r}
aic8=rep(0,2)
glm4=glm(form1pgpt,family = Gamma(identity),data=Film1,subset = -912)
glm4.s=step(glm4)
summary(glm4.s)
aic8[1]=AIC(glm4.s)
```

```{r}

glm4=glm(form2pgpt,family = Gamma(identity),data=Film1,subset = -912)
glm4.s.2=step(glm4)
summary(glm4.s.2)
aic8[2]=AIC(glm4.s.2)
```


```{r}
aic8 #form2 is best
plot(glm4.s.2)

cv.glm(Film1[-912,],glm4.s.2)$delta

```

open theatre resdual plots


LASSO
```{r}

library(glmnet)

Film1s<-Film1[,-c(1,3,4,5,13,14,15,16,17,18,19,20,21)]

X=model.matrix(~., Film1s)[,-1]

cv.glmn1l<-cv.glmnet(X,log(Film1$Opening.weekend))

glmn1l.1<-glmnet(X,log(Film1$Opening.weekend),lambda = 0.002308 )
coef(glmn1l.1)

res1=log(Film1$Opening.weekend)-predict(glmn1l.1,X)
p1=predict(glmn1l.1,X)
qqnorm(res)
qqline(res)

plot(predict(glmn1l.1,X),res)
lines(lowess(predict(glmn1l.1,X),res), col="red")

glmn1l.1s<-glmnet(X,log(Film1$Opening.weekend))
plot(glmn1l.1s,xvar="lambda")
abline(v=log(cv.glmn1l$lambda.min))

cv.glmn1l$cvm[which(min(cv.glmn1l$lambda))]


y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=p1,y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```

```{r}

Xi=model.matrix(~.^2, Film1s)[,-1]
cv.glmn1li<-cv.glmnet(Xi,log(Film1$Opening.weekend))

glmn1l.1i<-glmnet(Xi,log(Film1$Opening.weekend),lambda = 0.1766 )
coef(glmn1l.1i)

res1=log(Film1$Opening.weekend)-predict(glmn1l.1i,Xi)
p1=predict(glmn1l.1i,Xi)
qqnorm(res)
qqline(res)

plot(predict(glmn1l.1i,Xi),res1)
lines(lowess(predict(glmn1l.1i,Xi),res1), col="red")
#plot(glmn1l.1i)
#plot(glmn1l.1i,xvar = "lambda")
#plot(glmn1l.1i,xvar = "dev")
#plot(cv.glmn1li)

y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=p1,y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```

```{r}
Film1s2<-Film1[-912,-c(1,3,13,14,15,16,17,18,19,20,21)]

X2=model.matrix(~., Film1s2)[,-1]

cv.glmn2l<-cv.glmnet(X2,log(Film1$postopengross[-912]))

glmn2l.1<-glmnet(X2,log(Film1$postopengross[-912]),lambda = 0.001397 )
coef(glmn2l.1)

res1=log(Film1$postopengross[-912])-predict(glmn2l.1,X2)
p1=predict(glmn2l.1,X2)

qqnorm(res)
qqline(res)
plot(predict(glmn2l.1,X2),res)
lines(lowess(predict(glmn2l.1,X2),res), col="red")



y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=p1,y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```

```{r}
Film1s2<-Film1[-912,-c(1,3,13,14,15,16,17,18,19,20,21)]

X2i=model.matrix(~.^2, Film1s2)[,-1]

cv.glmn2li<-cv.glmnet(X2i,log(Film1$postopengross[-912]))

glmn2l.1i<-glmnet(X2i,log(Film1$postopengross[-912]),lambda = 0.002026 )
coef(glmn2l.1i)

res1=log(Film1$postopengross[-912])-predict(glmn2l.1i,X2i)
p1=predict(glmn2l.1i,X2i)

qqnorm(res)
qqline(res)
plot(predict(glmn2l.1i,X2i),res)
lines(lowess(predict(glmn2l.1i,X2i),res), col="red")


y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=p1,y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))


```


Regression Trees



```{r}
library(rpart)
library(randomForest)
library(partykit)
```

```{r}
form=as.formula(paste("log(Opening.weekend)~",paste(names(Film1)[c(2,6,7,8,9,10,11,12,15,16)],collapse = "+")))

rt1=rpart(form,Film1,method = "anova")
summary(rt1)

p1=predict(rt1)

res1=log(Film1$Opening.weekend)-p1
plot(jitter(p),res)
lines(lowess(p,res), col="red")

qqnorm(res)
qqline(res)

rt1.p=as.party(rt1)
plot(rt1.p)


y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=jitter(p1),y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```


```{r}
library(caret)
tc=trainControl(method="cv",number = 10)
cp=0.01
model=train(form,data=Film1,trControl=tc,method="rpart",tuneGrid=data.frame(cp))

```




```{r}
formp=as.formula(paste("log(postopengross)~",paste(names(Film1)[c(2,4,5,6,7,8,9,10,11,12,15,16)],collapse = "+")))

rt2=rpart(formp,Film1,subset = -912)
summary(rt2)
p1=predict(rt2)

res1=log(Film1$postopengross)[-912]-p1
plot(jitter(p),res)
lines(lowess(p,res), col="red")
qqnorm(res)
qqline(res)

rt2.p=as.party(rt2)
plot(rt2.p)

y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=jitter(p1),y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```

```{r}
cp=0.01
model=train(formp,data=Film1[-912,],trControl=tc,method="rpart",tuneGrid=data.frame(cp))

model=train(formp,data=Film1[-912,],trControl=tc,method="rpart")
```


Random Forest

```{r}
library(randomForest)

rf=randomForest(form,Film1,importance=T)

varImpPlot(rf)
res1=log(Film1$Opening.weekend)-rf$predicted
p1=rf$predicted
plot(rf$predicted,res)
lines(lowess(rf$predicted,res), col="red")
qqnorm(res)
qqline(res)
importance(rf)

y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=jitter(p1),y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))
```


```{r}
mtry=3
model=train(Film1[,c(2,6,7,8,9,10,11,12,15,16)],log(Film1$Opening.weekend),trControl=tc,method="rf",tuneGrid=data.frame(mtry))
#tuneGrid=data.frame(mtry)
```

```{r}
rf2=randomForest(formp,Film1,subset = -912,importance=T)

varImpPlot(rf2)
res1=log(Film1$postopengross)[-912]-rf2$predicted
p1=rf2$predicted
plot(rf2$predicted,res)
lines(lowess(rf2$predicted,res), col="red")
qqnorm(res)
qqline(res)
importance(rf2)


y=quantile(res1,c(0.25,0.75))
x=qnorm(c(0.25,0.75))
slope=diff(y)/diff(x)
int=y[1]-slope*x[1]

g=ggplot(as.data.frame(cbind(p1,res1)))
g+geom_point(aes(x=jitter(p1),y=res1))+geom_smooth(aes(x=p1,y=res1),method="loess",se=FALSE,col="blue")+xlab("Predicted values")+ylab("Residuals")
g+geom_abline(slope=slope,intercept = int,color="grey",linetype=2)+stat_qq(aes(sample=res1))

```

```{r}
mtry=4
model2=train(formp,data=Film1[-912,],trControl=tc,method="rf",tuneGrid=data.frame(mtry))
```

```{r}
res=log(Film1$postopengross)[-912]-model$finalModel$predicted
plot(model$finalModel$predicted,res)
```

